---
title: "Practical 2 - GLMM"
format: 
  html:
    theme:
      light: flatly
      dark: darkly
  PrettyPDF-pdf:
    keep-tex: true
    number-sections: true
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(webexercises)

```

In this practical we are going to fit a Generalized (Mixed) Linear Model in `inlabru`.

We are going to:

- Fit a Poisson regression
- Change the prior distributions for the model hyperparameters
- Compute and visualize posterior densities and summaries for marginal effects
 
 
 We start by loading some useful libraries 
 
```{r}
#| warning: false
#| message: false

library(dplyr)
library(INLA)
library(ggplot2)
library(patchwork)
library(inlabru)     
library(MASS)

```

In this practical we are going to analyse the dataset `eagles`. The data, contained  in the library `MASS`, are records of salmon pirating attempts by Bald Eagles in Washington State. See `?eagles` for details. 

While one eagle feeds, sometimes another will swoop in and try to steal the salmon from it. Call the feeding eagle the “victim” and
the thief the “pirate.”

```{r}
data(eagles)
?eagles
print(eagles)
```

## Fitting Poisson regression
```{r}
library(MASS)
data(eagles)
eagles = eagles %>% 
  mutate(P = ifelse(P == "L", 1,0),
         A = ifelse(A == "A", 1,0),
         V = ifelse(V == "L", 1,0),
         )
```


```{r}
cmp = ~ Intercept(1) + fixed(~ -1 + P + A + V, model = "fixed")
lik = bru_obs(formula = y ~ .,
              N  = n,
              family = "binomial",
              data = eagles)
fit1 = bru(cmp, lik)

cmp = ~ Intercept(1) + fixed( V, model = "iid", initial  = 0.1, fixed = T)



```




```{r}
cmp = ~ Intercept(1, prec.linear = 0.02) + fixed(~ -1 + P + A + V, model = "fixed", hyper = list(prec = list(initial = log(0.01), fixed = T)))
lik = bru_obs(formula = y ~ .,
              N  = n,
              family = "binomial",
              data = eagles)
fit2 = bru(cmp, lik)
```



```{r}

library(lme4)
data("grouseticks")

head(grouseticks)


grouseticks  = grouseticks %>%
  group_by(BROOD, YEAR) %>%
  mutate(ij = cur_group_id()) %>%
  ungroup() %>%
  mutate(ijk = seq_along(INDEX))

cmp= ~ -1 + year(YEAR, model = "iid", fixed = T, initial = log(0.01)) +
  height(HEIGHT, model =  "linear") + 
  brood_year(ij , model = "iid") +
  brood_year_chicken(ijk, model= "iid")

lik = bru_obs(TICKS ~ .,
              data = grouseticks,
              family = "Poisson")

  
fit = bru(cmp, lik)

```

 
